Spatial Graph-Based Social Trade\\ Transaction Simulation

### 2017-05-26 paper/paper.tex
Basic principles, minor changes

% EXTRA PACKAGES. Please place here any extra package you need along with your own command definitions

@@

Private variables:
\begin{description*}
	\item[resources]\text{}\\
		Variable of type \texttt{Double}, ranging from 0 to (potentially) infinity. This represents how much an agent possesses and influences his trading behaviour.

	\item[sympathy]\text{}\\
		Variable of type \texttt{int}, ranging from (potentially) minus infinity to (potentially) plus infinity. This represents how much an agent likes one specific neighbour. It influences its trading behaviour and is adjusted as explained below.

	\item[position on grid]\text{}\\
		The $x$- and $y$-coordinates (of type \texttt{int}) of an agent on the plane.
\end{description*}

Static variables:
\begin{description*}
	\item[willingness-to-trade-factor ($wtf$)]\text{}\\
		Variable of type \texttt{double}, ranging from 0 to 1. This represents the chance that an agent actually makes a trade with someone else, independently of all the other factors.

	\item[gold-dig-factor ($gdf$)]\text{}\\
		Variable of type \texttt{Double}, ranging from 0 to 1. This represents how an agent decides on his trading partner. If the value is close to 0, the agent will choose a partner which he likes, disregarding the amount of resources this potential partner has. If the value is close to 1, the agent will choose a partner which has lots of resources, disregarding how much he likes him.

	\item[circe-of-life-factor ($clf$)]\text{}\\
		Variable of type \texttt{Double}, ranging from 0 to 1. This represents the chance that an agent dies or spawns another agent (representing sexual reproduction) in any given iteration independent of other factors or variables.
\end{description*}

@@

To determine how the resources are distributed, we use a variable called \textbf{merciful-god-factor} ($mgf$) and a distribution function $f: \texttt{int} \times \texttt{int} \to \texttt{Double}$. If the $mgf$ is equal to 1, the sum of all resources dsitributed on the plane are sufficient for all agents to survive. If the $mgf$ is above 1, there is an excess of resources, if the $mgf$ is below 1, there is a shortage (linearly dependent on the $mgf$).

\subsection{Process overview and scheduling}
Our simulation works with synchronous updates (ticks), meaning all agents do the same things at the same time. The actions in each tick can be divided into three main parts:
\subsubsection{Resource distribution}
		    At the beginning of each tick the resources each parcel receives are calculated and distributed accordingly. Now all agents have the possibility to propose a trade with exactly one of their neighbours. They do this by finding the best match for a trading partner in the list of their neighbours. The best trading partner is found by computing
	$$v = gdf \cdot resources + (1-gdf) \cdot sympathy$$
for each neighbour and selecting the best candidate (candidate with the highest value $v$).

@@

\subsubsection{Basic Principles}
	social behaviour on a scale from only sympathy to only rationality (gold-dig-factor)

>>>>

% EXTRA PACKAGES. Please place here any extra package you need along with your own command definitions
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\argmax}{arg\,max}
\newcommand{\gdf}{\mathit{GDF}}
\newcommand{\wtf}{\mathit{WTF}}
\newcommand{\clf}{\mathit{CLF}}
\newcommand{\col}{\mathit{COL}}
\newcommand{\mgf}{\mathit{MGF}}

@@

Private variables:
\begin{description*}
	\item[Resources:]
		Variable ranging from 0 to (potentially) infinity. This represents how much an agent possesses and influences his trading behaviour.

	\item[Sympathy:]
		Variable which represents how much an agent likes one specific neighbour. It influences its trading behaviour and is adjusted as explained below.

	\item[Position on Grid:]
		The $x$- and $y$-coordinates of an agent on the plane. This only represents the spatial location, disregarding its position in the network.
\end{description*}

Static variables:
\begin{description*}
	\item[willingness-to-trade-factor ($\wtf$):]
		Percentage representing the chance that an agent actually makes a trade with someone else, independently of all the other factors.

	\item[gold-dig-factor ($\gdf$):]
		Ratio representing how an agent decides on his trading partner. It can be visualized with the following fraction:
		\begin{equation*}
			\gdf = \frac{\text{\textit{Sympathy based trading}}}{\text{\textit{Economics based trading}}}
		\end{equation*}
		If the value is close to 0, the agent will choose a partner which he likes, disregarding the amount of resources this potential partner has. If the value is close to 1, the agent will choose a partner which has lots of resources, disregarding how much he likes him.

	\item[circe-of-life-factor ($\clf$):]
		Percentage representing the chance that an agent dies or spawns another agent (representing sexual reproduction) in any given iteration independent of other factors or variables.

	\item[Cost of Life ($\col$):]
		The amount of resources which an agent has to spent in each tick. This value is also used to determine the distribution of resources.
\end{description*}

@@

To determine how the resources are distributed, we use a variable called \textbf{merciful-god-factor} ($\mgf$). If the $\mgf$ is equal to 1, the sum of all resources distributed on the plane are sufficient for all agents to survive. If the $\mgf$ is above 1, there is an excess of resources, if the $\mgf$ is below 1, there is a shortage (linearly dependent on the $\mgf$).

\subsection{Process overview and scheduling}
Our simulation works with synchronous updates (ticks), meaning all agents do the same things at the same time. The actions in each tick can be divided into three main parts:
\subsubsection{Resource distribution}
	To distribute the resources on the grid, we us the function
	\begin{equation}
		f_\mathit{dist}(\mathit{pos},\, \mathit{time}) \to [0,\, 1]
	\end{equation}
	which takes the position of a parcel as an input and calculates the percentage which determines how much of the total resources this specific parcel receives. To obtain the actual amount of resources distributed, we multiply this percentage with the total need for resources and the $\mgf$:
	\begin{equation}
		resource = f_\mathit{dist}(\mathit{pos},\, \mathit{time}) \cdot n \cdot \col \cdot \mgf
	\end{equation}
	where $n$ is the initial number of agents. Using this formula, we see that the total distribution of resources is equal to the total needs:
	\begin{alignat}{2}
		&\sum_{p \in \mathit{Parcels}} f_\mathit{dist}(\mathit{p.pos},\, \mathit{time}) \cdot n \cdot \col &&= n \cdot \col\\
		\Longleftrightarrow \quad &\sum_{p \in \mathit{Parcels}} f_\mathit{dist}(\mathit{p.pos},\, \mathit{time}) &&= 1
	\end{alignat}

\subsubsection{Trade requests}
	Now all agents have the possibility to propose a trade with exactly one of their neighbours. They do this by finding the best match for a trading partner in the list of their neighbours. The best trading partner is found by evaluating the following formula:
	\begin{equation}
		\argmax_{a \in \mathit{Neighbours}} \left( \gdf \cdot a.\mathit{resource} + (1-\gdf) \cdot \mathit{sympathy}(a) \right)
	\end{equation}
	where $sympathy()$ is the mapping from a neighbour of an agent to a sympathy value.

@@

\subsubsection{Basic Principles}
	In this section we will explain some of our design choices. The composition of these choices is what makes our model. First we will examine the individual decisions and explain the thinking behind them and after that we will explain why the composition of them is a useful approximation of the civilization we want to model. As mentioned in the introduction we will again distinguish and investigate along the axes of social behaviour and structure.

	social behaviour:
	\begin{description*}
		\item[$\gdf$:]
			The biggest part of our agent's characters is decided by the value of the gold-dig-factor (cf. gold-dig-factor in the section Entities, state variables and scales). The idea behind this is that a $\gdf$ of 1 causes behaviour similar to socialist/communist ideas while a $\gdf$ of 0 leads to phenomena similar to capitalism. Capitalism favours accumulation of money, be it in the hands of individuals or corporations [Karl Marx, Das Kapital, vol. 1, ch. 25]. On the other side of the spectrum socialism favours the equal distrubution of money as can be seen in socialist ideas such as progressive taxes [Moyes, P. A note on minimally progressive taxation and absolute income inequality]. We mirror this using the $\gdf$ in the following way: an agent with a high $\gdf$ will, when considering all agents that have proposed him a trade, favour the agent with the least resources over any other agent, no matter how much sympathy he has for any of them. It is obvious that this favours the equal distribution of resources since in a scenario of unequally distributed wealth with all agents having a $\gdf$ of 1 all poor agents will propose trades to (really it's requestint resources from) rich agents which, in turn will accept these trades without consideration. On the other hand a $\gdf$ of zero of near zero will favour the emergence of groups of agents which will exclusively help other agents of the group. Poor agents requesting resources will lose the sympathy of richer agents because requesting resources will lead to a decrease of the sympathy factor while rich agents who don't have to request resources very often will maintain a high sympathy factor.

		\item[resource:]
			The one resource of our model can be thought of as being money. Money can take have multiple functions: among others it is a medium of exchange, a measure of value and a store of value [William Stanley Jevons, Money and the Mechanism of Exchange]. In our model the agents need the resource (from now on denoted as money) in order to survive. It acts therefore as a medium of exchange, the agents can, hypothetically, go and buy food and whatever they need in order to survive. At the same time it is a store of value; as we'll see in the next section agents give away money but at the same time they obtain another abstract value, the sympathy of the receiving agent and therefore some form of insurance for the future. The value is not lost but stored in another form. Money also acts as a measure of value, it allows us to quantify the success of individual agents and groups of agents by assigning them and their position a value.

		\item[Trading:]
			As we've seen trading in our model is somehow unintuitive: agents can request a trade and then may or may not receive money without really giving something back. No trading actually takes place, it's more giving than trading. The assumption here is a creditor-debitor relationship. A rich agent will give money to a poor agent but both will remember that the trade has happened in the form of the changed sympathy. While the creditor is less likely to give to the same debitor many times, the debitor remembers his debt towards the creditor and is more likey to give him money in case of a request. This resembles the real world notion of credit in so far as the creditor is willing to wait for the debitors situation to improve before claiming back his money. How exactly this creditor-debitor relationship is to be interpreted depends again on the $\gdf$ of both participants: with a low $\gdf$ the actual effect of a trade is accurately described by the above, with a high $\gdf$ not so much since the selection of a trading partner depends less on the sympathy level. This is no contradiction though since a high $\gdf$ mirrors, as described before, the affinity of an agent to distribute wealth. Apart from the obvious distribution of resources the action of trading is an implicit channel of information between two agents. We will investigate this role of trade a bit more below in the section on the logical connection between agents.

		\item[Sympathy:]
			The total sympathy level is steadily rising as in each trade the net change of sympathy is +1 in contrast of the somewhat more natural approach of saying, the helper's sympathy towards the helped decreases and the helped agent's sympathy towards it's helper increases. Before explaining this we need to take a step back and explain the term. According to the Cambridge Academic Content Dictionary sympathy can mean two things: "a feeling or expression of understanding and caring for someone else who is suffering or has problems that have caused unhappiness" or "a feeling or expression of support and agreement". We use the latter definition for our model although the two are obviously intertwined in our model sympathy is not a reaction to the state of another agent but rather the state of the relationship to this actor. Among other things the sympathy one feels towards someone else is influenced by the proximity to this person, the similarity of one's state to this persons's state and "one's past and vicarious experiences" [Lowenstein, G.; Small, D. A. (2007). "The scarecrow and the tin man: The vicissitudes of human sympathy and caring"]. We express these factors in the way the sympathy factor changes, especially in the fact that the net sympathy is rising. Obviously the proximity between two individuals (thus, also between agents) increases with more interaction and since trade is The interaction in our model it surely should increase the sympathy. On the other hand an agent that has received help has found itself in a situation of need and is thus more likely to being able to relate to other agents being in such a state. 
	\end{description*}

	structure:
	\begin{description*}
		\item[Logical Connections:]
			The agents and the edges between them form a graph or a network. When there exists an edge between two agents it means that they "know" each other and that they are able to trade with each other. In this sense the network really captures the meaning of a social network, especially as we assume that all social interaction is trade. As shortly mentioned above the connection between two agents also acts as an implicit channel of information between them: an agent receiving a trade request from one of his neighbours learns something about this agent's situation and, as the situation of an agent mostly depends on his position on the map, also about the surrounding world. A well connected agent is thus better informed than others and is thus more likely to be able to adopt his position (cf: movement) better than others.

		\item[Phyisical Connection:]
			Agents which are close to each other are not necessarily logically connected by an edge, nevertheless they share a common piece of information in the shape of their parcel. Close neighbours are likely to be situated on the same parcel, thereby receiving the same amount of money and having a similar situation. As in the real world the milieu is somehow related to the wellbeing of its inhabitants. Obviously the physical position of an agent is related to its logical connections. An agent won't, for example, maintain relationships with agents which are very far away for a long time. We ensure this by periodiaclly updating the logical network of an agent to favour closer agents to agents which are further away.

		\item[Movement:]
			The last piece of the social structure is the ability of agents to change their position and thus change the structure of model. Although the only thing that changes is the position of the agent on the grid and not its logical place in the network. 
	\end{description*}
